---
title: "time_series"
output: html_document
date: "2024-10-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Loading and Processing

```{r loading packages}
# install.packages("devtools")
devtools::install_github("logan-berner/LandsatTS", build_vignettes = TRUE)
reticulate::use_virtualenv("rgee", required = TRUE)

# install.packages("rgee")
library(rgee)
# Recreate the Python environment with all dependencies
ee_install()

ee_check()

ee_Initialize()
```

```{r}

# Load packages for data handling etc.
library(sf)
library(dplyr)
library(purrr)
library(data.table)
library(stringr)
library(rgee)

# Load LandsatTS package
library(LandsatTS)

# Intialize the Earth Engine with rgee
ee_clean_user_credentials()
ee_Authenticate()
ee_Initialize(email = "lisamliubovich@gmail.com", drive = TRUE)
```

```{r loading globe data}
library(tidyverse)
globe_c <- read_csv("/Users/lisaliubovich/Desktop/grad_school/fall_2024/data-793/AoP/data:eda/globe_cv2.csv")
```

```{r subsetting the measured lat and long columns}
library(sf)
# Convert the coordinates (measured_lat, measured_long) to an sf object
globe_sf <- globe_c %>%
  st_as_sf(coords = c("measure_long", "measure_lat"), crs = 4326) %>%
  mutate(sample_id = row_number())  # Assign a unique sample ID to each point

# View the sf object
print(globe_sf)
```

```{r}
# Re-export the Landsat time-series data with fresh IDs
# task_list <- lsat_export_ts(globe_sf)

# Start the new tasks
# lapply(task_list, function(task) ee$batch$Task$start(task))

# Monitor the new tasks
# map(task_list, ee_monitoring)

```

```{r converting into satellite data}
# Load necessary libraries
library(data.table)  # For fast reading and handling of large datasets
library(dplyr)       # For data manipulation

# Set the path to the folder containing your exported files
export_folder <- "~/Desktop/grad_school/fall_2024/data-793/AoP/data:eda"

# List all CSV files in the folder
csv_files <- list.files(path = export_folder, pattern = "*.csv", full.names = TRUE)

# Function to read each file and ensure all columns are converted to character type
read_as_characters <- function(file) {
  data <- fread(file)
  data[] <- lapply(data, as.character)  # Convert all columns to character
  return(data)
}

# Use the custom function to read and combine all CSV files
lsat_data <- rbindlist(lapply(csv_files, read_as_characters), fill = TRUE)

# Check the data to ensure it loaded correctly
head(lsat_data)

```

```{r merge ground and satellite data}

lsat_data <- lsat_data %>%
  select(
    organization_id, org_name, site_id, site_name, latitude, longitude, elevation, measured_on, 
    landcover_id, data_source, measured_at, muc_code, muc_description, north_photo_url, 
    east_photo_url, south_photo_url, west_photo_url, upward_photo_url, downward_photo_url, 
    measure_lat, measure_long, measure_elev, loc_method, loc_accuracy, snow_ice, standing_water, 
    muddy, dry_ground, leaves_on_trees, raining_snowing
  ) %>%
  mutate(
    snow_ice = ifelse(!is.na(snow_ice), 0, 1),
    standing_water = ifelse(!is.na(standing_water), 0, 1),
    muddy = ifelse(!is.na(muddy), 0, 1),
    dry_ground = ifelse(!is.na(dry_ground), 0, 1),
    leaves_on_trees = ifelse(!is.na(leaves_on_trees), 0, 1),
    raining_snowing = ifelse(!is.na(raining_snowing), 0, 1)
  ) %>%
  filter(if_all(everything(), ~ !is.na(.)))

library(sf)
# Convert the coordinates (measured_lat, measured_long) to an sf object
lsat_data <- lsat_data %>%
  st_as_sf(coords = c("measure_long", "measure_lat"), crs = 4326) %>%
  mutate(sample_id = row_number())  # Assign a unique sample ID to each point

# View the sf object
print(lsat_data)

# Perform spatial join using st_join
merged_data <- st_join(lsat_data, globe_sf)

# Check the merged data
head(merged_data)

```

# Comparing Similarities and Differences

```{r comparing surface conditions}
# Calculate the agreement between snow_ice observations from satellite (.x) and ground (.y)
surface_conditions <- merged_data %>%
  summarise(
    snow_ice_agreement = mean(snow_ice.x == snow_ice.y, na.rm = TRUE),
    muddy_agreement = mean(muddy.x == muddy.y, na.rm = TRUE),  # Replace 'muddy' with actual column names
    standing_water_agreement = mean(standing_water.x == standing_water.y, na.rm = TRUE)  # Replace with actual column names
  )

print(surface_conditions)

```

```{r}
# Convert the spatial object back to a data frame to remove the geometry
merged_data_df <- as.data.frame(merged_data)

# Now summarize or manipulate the data as a regular data frame
surface_conditions_summary <- merged_data_df %>%
  summarise(
    snow_ice_agreement = mean(snow_ice.x == snow_ice.y, na.rm = TRUE),
    muddy_agreement = mean(muddy.x == muddy.y, na.rm = TRUE),
    standing_water_agreement = mean(standing_water.x == standing_water.y, na.rm = TRUE)
  )

print(surface_conditions_summary)



library(ggplot2)
# Visualize the snow_ice agreement distribution
ggplot(merged_data_df, aes(x = snow_ice.x, fill = factor(snow_ice.y))) +
  geom_bar(position = "dodge") +
  labs(title = "Comparison of Snow/Ice Conditions: Satellite vs. Ground", x = "Satellite Snow/Ice", fill = "Ground Snow/Ice")


```

-   **Agreement = 1**: This means that for the given data points, the satellite and ground observations match exactly for `snow_ice`, `muddy`, and `standing_water`.

-   **Geometry**: The result is a `MULTIPOINT` feature, meaning that this is still a spatial object containing the geometry (latitude/longitude coordinates) of the data points.

```{r}
# Convert elevation columns to numeric
merged_data <- merged_data %>%
  mutate(elevation.x = as.numeric(elevation.x),
         elevation.y = as.numeric(elevation.y))
# Check for non-numeric values (NA after conversion)
sum(is.na(merged_data$elevation.x))
sum(is.na(merged_data$elevation.y))


# Calculate the difference in elevation between GO points and satellite measurements
merged_data <- merged_data %>%
  mutate(elevation_diff = abs(elevation.x - elevation.y))  # Adjust column names accordingly

# Summarize the elevation differences
elevation_summary <- merged_data %>%
  summarise(mean_elevation_diff = mean(elevation_diff, na.rm = TRUE),
            max_elevation_diff = max(elevation_diff, na.rm = TRUE),
            min_elevation_diff = min(elevation_diff, na.rm = TRUE))

print(elevation_summary)
```

-   **Mean Elevation Difference**: 0.0408

-   **Max Elevation Difference**: 8.5

-   **Min Elevation Difference**:0

This suggests that, on average, the elevation between the satellite and ground measurements is quite close, with the largest difference being 8.5 units (probably meters)

```{r}
library(ggplot2)

# Plot elevation differences
ggplot(merged_data, aes(x = elevation.x, y = elevation.y)) +  # Replace with actual column names
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Comparison of Elevation: GO Points vs Satellite", 
       x = "GO Point Elevation", 
       y = "Satellite Elevation")


```

-   Each point on the scatter plot corresponds to an observation where elevation data were available from both the GO points and the satellite.

-   The **red dashed line** represents the **ideal line of agreement** where the satellite elevation perfectly matches the GO point elevation (i.e., `Satellite Elevation = GO Point Elevation`).

-   The points mostly fall along the red dashed line, indicating a high level of agreement between the satellite and GO point elevation measurements.

<!-- -->

-   If a point lies exactly on the line, it means that the **satellite and ground (GO point) elevation values are the same** for that observation.

-   Deviations from the line (if any) would indicate differences between the satellite and ground elevation measurements. However, from the plot, the points seem to be very close to the line, suggesting **minimal differences** between the two datasets.
